Classification of Histopathological Images of Breast Cells between Malignant and Benign Tumors based on Handcrafted Feature Extraction

### 2019-11-22 docs/article.tex
Some corrections

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.7\linewidth]{images/fourrierFiltering/maskedImage}
		\caption{Image with mask}
		\label{fig:maskedimage}
	\end{figure}
\section{Segmentation}
\par The segmentation is an important step within the image processing, which is responsible for subdividing the an image in regions or objects that compose the signal \cite{Gonzalez:2006:DIP:1076432}, usually, this process is made through the separation of the image between background and object. Several techniques can be used in this process, such as based on thresholding, edge detection, region approaches. In this work, the optimal thresholding technique called Otsu was chosen due to its simplicity and good results. The segmentation was made after the filtering step, and the approach taken in the present work is to subtract the background of the segmented image from the original gray levels image. The code responsible for doing this task and the result produced in the image are shown in the Figure 4.

@@

\section{Feature Extraction}
The Feature Extraction can be considered as one of the most important steps in Image Processing: this is the last phase before the classification of the samples. So, the main objective here is to compose feature vectors capable to describe the samples, in order to classify correctly each example in its real class.\par
Commonly, two approaches can be chosen in order to extract descriptors from an image: Handcrafted Features and Non-Handcrafted ones. The Handcrafted Features, as the name claims, are the descriptors extracted manually, according to choice of the developer, while the Non-Handcrafted Features are extracted in an automatic way, in order to do the best possible classification (one example of this approach are the Convolutional Neural Networks - CNN). \par
As mentioned in the title of this paper, the used approach for extracting the descriptors from the histopathological images was the handcrafted one, where 11 descriptors were extracted and used as the input of the classifiers. It is important to remember that the extraction was accomplished from segmented images and non-segmented ones, so, it is possible to compare the impact of the segmentation step during the classification. Below is presented the extracted features used in the process.\par
The descriptors used in the present process can be divided into 3 categories: First Order Measures, Second Order Measures and Others. The First Order Measures are the descriptors that consider just the gray level information about the pixels, not considering the spatial relation among them. The Second Order Measures, instead of the first ones, consider the spatial relation among the pixels, beyond the gray level property. These descriptors are extracted from the called Gray Level Co-occurrence Matrix (GLCM), that is responsible to store the frequency that a pixel with gray level \textbf{i} and other with gray level \textbf{j} occur in the image, separated for a distance \textbf{d}, in the direction\textbf{ $\theta$} \cite{conci2008computacao}. This method was proposed in 1973 by \cite{haralick:smc1973}, and the descriptors extracted from the GLCM are called Haralick descriptors (in the present work 6 of the 14 descriptors proposed by Haralick were used).\par
The Others Group mentioned in the paragraph above includes the descriptors that do not fit the First and Second Measures, so, becomes important to explain how these descriptors were extracted in the proposed method. Three descriptors can be found in this last group: Fourier Descriptors, Local Binary Pattern Features and Fractal Dimension Descriptors. \par
The Fourier Descriptors (FD), as the name say, are the features extracted from the Fourier Transform, that is responsible to transform the visualization of the signal from the spatial domain to the frequency domain, in order to become possible to see the frequencies that compose the image. Different approaches can be chosen to extract the FD`s, in the present work the proposed process was as follows:
\begin{itemize}
    \item The image was transformed to the frequency domain with the Fast Fourier Transform (FFT).
    \item Without doing the shift of the transformed, a number of N coefficients in the middle of the spectrum were taken, so just the high frequencies was caught from the FFT. Doing this process, it is possible to reconstruct the image just with these high frequencies coefficients, that represent the edges inside the signal. In the Figure 8 it is possible to see the original image and the reconstructed one made from the descriptors.
    \item Lastly, the complex magnitude was computed for the coefficients.
\end{itemize}{}

\begin{figure}[h]
    \centering
    \includegraphics[width=8cm]{images/imagem_reconstrucao.png}
    \caption{Original Image and its reconstruction using the Fourier Descriptors}
    \label{fig:my_label}
\end{figure}{}
The Local Binary Pattern Features (LBP) was described in \cite{lbp-OjalaPH96}, being a two-level version of the Texture Unit model proposed by \cite{Wang:1990:TCU:83597.83606}. The main idea is describe the image considering a neighborhood of 3x3 pixels and comparing the gray levels.\par
The Fractal Dimension Descriptors extracted in the present work were obtained using the Box Counting Theorem, a method where the image is divided into N windows second a certain scale and the number of boxes that contains objects are counted. Doing this, it is possible to obtain the Fractal Dimension from each scale, as well as a General one calculated through the angular coefficient.\par
The Figure 9 summarizes the descriptors obtained in the present work. 
\begin{figure}[h]
    \centering
    \includegraphics[width=8.5cm]{images/table_descriptors.png}
    \caption{Descriptors}
    \label{fig:my_label}
\end{figure}{}

@@

\section{Classification}
The relevance of the descriptors, as well the performance of the classifiers were measured by the analysis of ROC Curve and area under curve ROC (AUC). This measures were used in the proposed method because the malignant an benign classes are unbalanced in dataset, being that there are 24 malignant samples and 26 benign ones. Taking this is consideration, it is not a good idea analysing the obtained results through measures as accuracy, because the calculation of this last measure does not take the class balancing into consideration.\par
The ROC Curve and AUC allow to analyse the performance of the classifiers through the confusion matrix obtained varying the threshold, so the plotted graph is formed by the X axis (1 - Specificity) and Y (Sensitivity), therefore, the AUC with value equals to 1 represents a perfect classification, while an AUC equals to 0.5 or less represents a random classification.\par
The performance of the proposed method was evaluated through the classifiers \textbf{Discriminant Analysis}, \textbf{Decision Tree} and \textbf{Naive Bayes}, taking into consideration that all of these classifiers were submitted to the K-Folds validation method, with K equals to 5.\par
The best number of descriptors found by testing the classifiers used in the present method can be seen in the Figure 10, which shows the number of descriptors used for the classification in X axis, and the average rates for AUC in the Y axis.\par
It is important to remember that the graph contained in the Figure 10 was obtained using the segmented images in the process; the best number of descriptors obtained (35) was also tested in the same process without the segmentation step, in order to determine the importance of this phase in the whole process. The AUC rates obtained for the Discriminant Analysis, Decision Tree and Naive Bayes were 0.83, 0.66 and 0.75 respectively, resulting in an average of 0.746, well below the value showed in the graph (0.823). This information show us that the segmentation step produces a significant improvement in the performance of the classifiers, being necessary use its during the proposed method.

@@

\section{Results}

discrim_analysis
decision_tree
naive_bayes

>>>>

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.7\linewidth]{images/fourrierFiltering/maskedImage}
		\caption{Image with mask}
		\label{fig:maskedimage}
	\end{figure}
\section{Segmentation}
\par The segmentation is an important step within the image processing, which is responsible for subdividing  an image in regions or objects that compose the signal \cite{Gonzalez:2006:DIP:1076432}, usually, this process is made through the separation of the image between background and object. Several techniques can be used in this process, such as based on thresholding, edge detection, region approaches. In this work, the optimal thresholding technique called Otsu was chosen due to its simplicity and good results. The segmentation was made after the filtering step, and the approach taken in the present work is to subtract the background of the segmented image from the original gray levels image. The code responsible for doing this task and the result produced in the image are shown in the Figure 4.

@@

\section{Feature Extraction}
The Feature Extraction can be considered as one of the most important steps in Image Processing: this is the last phase before the classification of the samples. So, the main objective here is to compose feature vectors capable to describe the samples, in order to classify correctly each example in its real class.\par
Commonly, two approaches can be chosen in order to extract descriptors from an image: Handcrafted Features and Non-Handcrafted ones. The Handcrafted Features, as the name claims, are the descriptors extracted manually, according to choice of the developer, while the Non-Handcrafted Features are extracted in an automatic way, in order to do the best possible classification (one example of this approach are the Convolutional Neural Networks - CNN). \par
As mentioned in the title of this paper, the used approach for extracting the descriptors from the histopathological images was the handcrafted one, where 11 descriptors were extracted and used as the input of the classifiers. It is important to remember that the extraction was accomplished from segmented images and non-segmented ones, so, it is possible to compare the impact of the segmentation step during the classification. Below is presented the extracted features used in the process.\par
The descriptors used in the present process can be divided into 3 categories: First Order Measures, Second Order Measures and Others. The First Order Measures are the descriptors that consider just the gray level information about the pixels, not considering the spatial relation among them. The Second Order Measures, instead of the first ones, consider the spatial relation among the pixels, beyond the gray level property. These descriptors are extracted from the called Gray Level Co-occurrence Matrix (GLCM), that is responsible to store the frequency that a pixel with gray level \textbf{i} and other with gray level \textbf{j} occur in the image, separated for a distance \textbf{d}, in the direction\textbf{ $\theta$} \cite{conci2008computacao}. This method was proposed in 1973 by \cite{haralick:smc1973}, and the descriptors extracted from the GLCM are called Haralick descriptors (in the present work 6 of the 14 descriptors proposed by Haralick were used).\par
The ther group mentioned in the paragraph above includes the descriptors that do not fit the First and Second Measures, so, becomes important to explain how these descriptors were extracted in the proposed method. Three descriptors can be found in this last group: Fourier Descriptors, Local Binary Pattern Features and Fractal Dimension Descriptors. \par
The Fourier Descriptors (FD), as the name say, are the features extracted from the Fourier Transform, that is responsible to transform the visualization of the signal from the spatial domain to the frequency domain, in order to become possible to see the frequencies that compose the image. Different approaches can be chosen to extract the FD`s, in the present work the proposed process was as follows:
\begin{itemize}
    \item The image was transformed to the frequency domain with the Fast Fourier Transform (FFT).
    \item Without doing the shift of the transformed, a number of N coefficients in the middle of the spectrum were taken, so just the high frequencies was caught from the FFT. Doing this process, it is possible to reconstruct the image just with these high frequencies coefficients, that represent the edges inside the signal. In the Figure 8 it is possible to see the original image and the reconstructed one made from the descriptors.
    \item Lastly, the complex magnitude was computed for the coefficients.
\end{itemize}{}

\begin{figure}[h]
    \centering
    \includegraphics[width=8cm]{images/imagem_reconstrucao.png}
    \caption{Original Image and its reconstruction using the Fourier Descriptors}
    \label{fig:my_label}
\end{figure}{}
The Local Binary Pattern Features (LBP) was described in \cite{lbp-OjalaPH96}, being a two-level version of the Texture Unit model proposed by \cite{Wang:1990:TCU:83597.83606}. The main idea is describe the image considering a neighborhood of 3x3 pixels and comparing the gray levels.\par
The Fractal Dimension Descriptors extracted in the present work were obtained using the Box Counting Theorem, a method where the image is divided into N windows according to a certain scale and the number of boxes that contains the objects that must be counted. Doing this, it is possible to obtain the Fractal Dimension from each scale, as well as a General one calculated through the angular coefficient.\par
The Figure 9 summarizes the descriptors obtained in the present work. 
\begin{figure}[h]
    \centering
    \includegraphics[width=8.5cm]{images/table_descriptors.png}
    \caption{Descriptors}
    \label{fig:my_label}
\end{figure}{}

@@

\section{Classification}
The relevance of the descriptors, as well the performance of the classifiers were measured by the analysis of ROC Curve and area under curve ROC (AUC). This measures were used in the proposed method because the malignant an benign classes are unbalanced in dataset, considering that there are 24 malignant samples and 26 benign ones. Taking this is consideration, it is not a good idea analysing the obtained results through measures as accuracy, because the calculation of this last measure does not take the class balancing into consideration.\par
The ROC Curve and AUC allow to analyse the performance of the classifiers through the confusion matrix obtained varying the threshold, so the plotted graph is formed by the X axis (1 - Specificity) and Y (Sensitivity), therefore, the AUC with value equals to 1 represents a perfect classification, while an AUC equals to 0.5 or less represents a random classification.\par
The performance of the proposed method was evaluated through the classifiers \textbf{Discriminant Analysis}, \textbf{Decision Tree} and \textbf{Naive Bayes}, taking into consideration that all of these classifiers were submitted to the K-Folds validation method, with K equals to 5.\par
The best number of descriptors found by testing the classifiers used in the present method can be seen in the Figure 10, which shows the number of descriptors used for the classification in X axis, and the average rates for AUC in the Y axis.\par
It is important to remember that the graph contained in the Figure 10 was obtained using the segmented images in the process; the best number of descriptors obtained (35) was also tested in the same process without the segmentation step, in order to determine the importance of this phase in the whole process. The AUC rates obtained for the Discriminant Analysis, Decision Tree and Naive Bayes were 0.83, 0.66 and 0.75 respectively, resulting in an average of 0.746, well below the value showed in the graph (0.823). This information show us that the segmentation step produces a significant improvement in the performance of the classifiers, being necessary use its during the proposed method.

@@

\section{Results}
	\par The classifiers used 
discrim analysis
decision tree
naive bayes
